Dependently Typed Languages in Statix

### 2023-04-24 pres-thesis/presentation.tex
Minor changes based on more practice runs
typo
Changes based on meeting

\begin{exampleblock}{Example}
\begin{lstlisting}
int x = 5;
int y = x + 3;
return x + y;
\end{lstlisting}
\end{exampleblock}

@@

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
\begin{itemize}
	\item Will it be easier than doing it in Haskell?
	\item Defining system F was a challenge\footnote{Hendrik van Antwerpen et al. Scopes as types.}, will it even be possible?
\end{itemize}

@@

	\begin{exampleblock}{Example}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (T: Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}

@@

\begin{frame}[fragile]{Type Checking}
	\begin{figure}
		\includegraphics[width=0.9\linewidth]{img/screenshot007}
	\end{figure}
	% We define the language using inference rules
	% How can we express those rules in Statix?
\end{frame}

@@

	\begin{figure}
		\includegraphics[width=0.7\linewidth]{screenshot004}
	\end{figure}

@@

\begin{frame}[fragile]{Type Checking: Requires Evaluation}
	\begin{exampleblock}{Example from earlier}
		\texttt{let T = if false then Int else Bool end;\\let b: T = true;}
	\end{exampleblock}
	\begin{block}{Evaluation relation} 
		\texttt{betaHeadReduce : scope * Expr -> scope * Expr} \\
		\texttt{betaReduce : scope * Expr -> Expr} \\
		\texttt{exectBetaEq : (scope * Expr) * (scope * Expr)}
	\end{block}
\end{frame}

\begin{frame}[fragile]{Type Checking: Requires Evaluation}
	\begin{figure}
		\includegraphics[width=0.9\linewidth]{screenshot001}
	\end{figure}

@@

\begin{frame}{Comparison with Haskell}
	\begin{block}{Type checking}
		\texttt{data EnvEntry = NType Expr | NSubst SExpr
			\\type Env = [EnvEntry]
			\\type SExpr = (Env , Expr)
			\\tc :: SExpr -> Either String Expr}
	\end{block}

	\begin{block}{Differences}
		\begin{itemize}
			\item Distribution of definitions over files
			\item De Bruijn Indices vs Names
			\item Inference built-in to Statix
		\end{itemize}
	\end{block}
\end{frame}

@@

@@

\begin{frame}[fragile]{Conclusions}
	Spoofax is a great tool for developing dependently typed languages!
	\begin{itemize}
		\item We can use scopes to represent environments and contexts
		\item Statix can still use improvements
		% - Higher order 
	\end{itemize}

	%- using scopes as env+context  
	%- statix enforces syntactic uniqueness of metavariable solutions but dependently typed languages use a weaker notion of equality 
	%- full thesis coming soonTM
\end{frame}

% block
% exampleblock
% alertblock

>>>>

\begin{exampleblock}{Example}
\begin{lstlisting}
let x = 5;
let y = x + 3;
return x + y;
\end{lstlisting}
\end{exampleblock}

@@

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
\begin{itemize}
	\item Will it be easier than doing it in Haskell?
	\item Prior work: Defining System F\footnote{Hendrik van Antwerpen et al. Scopes as types.}
\end{itemize}

@@

	\begin{exampleblock}{Example}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}

@@

@@

	\begin{figure}
		\includegraphics[width=0.7\linewidth]{screenshot005}
	\end{figure}

@@

\begin{frame}[fragile]{Type Checking: Requires Evaluation}
	\begin{exampleblock}{Example from earlier}
		\texttt{let T = if false then Int else Bool end;\\let b: T = true;}
	\end{exampleblock}
	\begin{block}{Evaluation relations} 
		\texttt{betaHeadReduce : scope * Expr -> scope * Expr} \\
		\texttt{betaReduce : scope * Expr -> Expr} \\
		\texttt{expectBetaEq : (scope * Expr) * (scope * Expr)}
	\end{block}

@@

@@

\begin{frame}{Comparison with Haskell}
	\begin{block}{Type checking}
		\texttt{data EnvEntry = NType Expr | NSubst (Env, Expr)
			\\type Env = [EnvEntry]
			\\tc :: Env -> Expr -> Either String Expr}
	\end{block}

	\begin{block}{Differences}
		\begin{itemize}
			\item Distribution of definitions over files
			\item De Bruijn Indices vs Names
			\item Definition in Statix is approximately 20\% longer
			\item Inference built-in to Statix
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}[fragile]{Conclusions}
	\begin{itemize}
		\item \textbf{Spoofax perspective:} Support for complex inference could be improved, for example using equational unification.
		\item \textbf{Dependent types perspective:} This already functions as a great tool for prototyping dependently typed languages!
		% - Higher order 
	\end{itemize}
\end{frame}
